Survival Analysis

### 2014-04-01 presentation.Rnw
Add formulas, make class activity
Fix up kaplan-meier slide, add n_i formula
Fixed desc. formatting, removed old stuff

\begin{tabular}{lllcc}
Survival
  & $P(\mbox{no event before } t)$
  & $S(t)$ & = & $1 - \int^{t}_{0} f(x)$ \\
Hazard
  & $OR(\mbox{no event until } t)$
  & $\lambda(t)$ & = & $^{f(t)} / _{S(t)}$ \\
Cumulative \\ Hazard
  & $OR(\mbox{event before } t)$
  & $\Lambda(t)$ & = & $\int^{t}_{0} \lambda(t) dt$ \\
\end{tabular}
\end{frame}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Old Stuff 
%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Introduction
\begin{frame}{Sounds cool\ldots but what is it?}
  \begin{itemize}
    \item What proportion will ``survive'' past a certain time?
    % Event History Analysis (sociology):
    %   Which kind of strikes last more than 3 days?
    \item Of those that ``survive,'' what is their failure or death rate?
    % Reliability Analysis (engineering):
    %   How long do different types of F1 tyres last?
    \item What factors affect how individuals ``survive''?
    % Duration Modeling (economics):
    %   How do different factors affect unemployment duration?
  \end{itemize}

  % also: political careers, earthquakes, medical, 
\end{frame}

%%% History
%%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
\begin{frame}{How did this magic happen?}
  \begin{itemize}
    \item Analyze actuarial data
    % for mortality tables, very informal
    \item Became a discipline in WWII
    % interest reliability of equipment
    % spurred on by spike in demands for reliability, quality
    \item Medical field drove much of survival analysis
    % b/c studies do not need all subjects at the start of the study (async)
    %   ...handles subjects that quit, cannot follow-up, or die of different event
    %   (handles incomplete data)
    \item Kicked off with Kaplan-Meier estimator
    % http://www.jehps.net/juin2009/Aalenetal.pdf
    % raises new questions: how to compare survival curves?
    % prop hazards resolved covariates problem
    % application was ahead of theory, didn't know why things worked
  \end{itemize}
\end{frame}
\end{document}

>>>>

\renewcommand{\arraystretch}{1.5}
\begin{tabular}{lllcc}
Survival
  & $P(\mbox{no event before } t)$
  & $S(t)$ & = & $1 - \int^{t}_{0} f(x)$ \\
Hazard
  & $OR(\mbox{no event until } t)$
  & $\lambda(t)$ & = & $^{f(t)} / _{S(t)}$ \\
Cum. Haz.
  & $OR(\mbox{event before } t)$
  & $\Lambda(t)$ & = & $\int^{t}_{0} \lambda(t) dt$ \\
  & & & = & $- \log{S(t)}$ \\
\end{tabular}
\end{frame}

@@

\begin{frame}{Estimators}
Kaplan-Meier Max Likelihood Estimator for $\hat S(t)$

\vspace{0.2in}
\begin{tabular}{ll}
  $t_i$ & time interval \\
  $d_i$ & \# events in $t_i$ \\
  $n_i$ & \# subjects \emph{at risk} in $t_i$ \\
\end{tabular}

\begin{eqnarray*}
  \hat S(t_0) & = & 1 \\
  \hat S(t)   & = & \prod_{t_i < t} \frac{n_i - d_i}{n_i} \\
              & = & \frac{n_i - d_i}{n_i} \cdot S(t_{i - 1}) \\ \\
    n_{i + 1} & = & n_i - d_i - c_i
\end{eqnarray*}
\end{frame}

@@

%%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
%%% http://www.jehps.net/juin2009/Aalenetal.pdf
\end{document}
